---
title: Continent COVID-19 LMIC Reports"
header-includes: \usepackage{caption}
date: ""
output:
  html_document:
    keep_md: no
    self_contained: yes
    css: styles.css
    theme: cosmo
    fig_caption: TRUE
    includes:
      in_header: ganalytics.html
  pdf_document: 
    fig_caption: true
    includes: 
      before_body: footer.tex
  word_document:
    df_print: kable
    fig_caption: yes
    fig_height: 4
    fig_width: 6
    highlight: tango
    reference_docx: template.docx
---

```{js, echo = FALSE}

if ($(window).width() < 768) {
  $('.dropdown-menu a.dropdown-toggle').on('click', function(e) {
    if (!$(this).next().hasClass('show')) {
      $(this).parents('.dropdown-menu').first().find('.show').removeClass("show");
    }

 

    var $subMenu = $(this).next(".dropdown-menu");
    if (!$subMenu.hasClass('show')) {
      $subMenu.addClass('show');
      $subMenu.show();
    } else {
      $subMenu.removeClass('show');
      $subMenu.hide();
    }

 

    $(this).parents('li.nav-item.dropdown.show').on('hidden.bs.dropdown', function(e) {
      $('.dropdown-submenu .show').removeClass("show");
    });

 

    return false;
  });
}

```

```{r, echo = FALSE, message=FALSE, warning=FALSE}
if (!exists("date")) { # running outside of orderly
  date <- Sys.Date()
}

data_url <- paste0("https://github.com/mrc-ide/global-lmic-reports/raw/master/data/",date,".csv")
url <- paste0("https://github.com/mrc-ide/global-lmic-reports/raw/master/",continent,"/index.pdf")

library(dplyr)
library(ggplot2)
library(squire)

format <- knitr::opts_knit$get("rmarkdown.pandoc.to")
figRef <- function(caption) {
  if(format == "latex") {
    gsub("Figure \\d: ","",caption)
  } else {
    caption
  }
}


fig_height <- function(pdf = 4, html = 6) {
    if(format == "latex") {
    pdf
  } else {
    html
  }
}

fig_center <- function() {
    if(format == "latex") {
    "default"
  } else {
    "center"
  }
}

latex_omit <- function(text) {
  if(knitr::is_latex_output())
    return("")
  else
    return(text)
}

pagebreak <- function() {
  if(knitr::is_latex_output())
    return("\\newpage")
  else
    return('<div style="page-break-before: always;" />')
}

linebreak <- function() {
  if(knitr::is_latex_output())
    return("")
  else
    return('<br>')
}

pdf_line_break <- function() {
  if(knitr::is_latex_output())
    return("<br>")
  else
    return('')
}


```

```{r ecdc prep, echo = FALSE, collapse=TRUE, warning=FALSE}

# get the ecdc data
df <- readRDS("ecdc_all.rds")
data <- df[,c("dateRep", "deaths", "cases","Region", "countryterritoryCode")]
names(data)[c(1,4,5)] <- c("date","country","iso3c")
data$date <- as.Date(data$date)
date <- date_0 <- as.Date(date)

# group for cumulative deaths
data <- group_by(data, country) %>% mutate(cum_deaths = rev(cumsum(rev(deaths))),
                                          cum_cases = rev(cumsum(rev(cases))),)



d <- data
d$country[d$country=="Congo"] <- "Republic of Congo"
d$country[d$country=="United_Republic_of_Tanzania"] <- "Tanzania"
d$country[d$country=="Curaçao"] <- "Curacao"
suppressWarnings(d$Continent <- countrycode::countrycode(d$country, origin = 'country.name', destination = 'continent'))
d$Continent[d$country=="Eswatini"] <- "Africa"
d$Continent[d$country=="United State of America"] <- "Americas"
d$Continent[d$country=="Isle_of_Man"] <- "Europe"             
d$Continent[d$country=="Kosovo"] <- "Europe"                  
d$Continent[d$country=="Netherlands_Antilles"] <- "Americas"    
d$Continent[d$country=="Saint_Lucia"] <- "Americas"             
d$Continent[d$country=="South_Korea"] <- "Asia"             
d$Continent[d$country=="United_States_of_America"] <- "Americas"
d$Continent[d$country=="Kosovo"] <- "Europe"

cont_sums <- d[d$Continent == continent & d$date == date_0,]

## summaries
summaries <- readRDS("summaries.rds")

```

## Situation Report for COVID-19: `r continent`, `r format(date, "%Y-%m-%d")`

`r latex_omit("---")`

#### **[Download the report for `r continent`, `r format(date, "%Y-%m-%d")` here](`r url`).**

<br> 

**Please see the country reports for further details on individual countries. For further guidance and questions about the methodology and caveats, please see the [Frequently Asked Questions](https://mrc-ide.github.io/global-lmic-reports/FAQ.html)**

<br>

### Epidemiological Situation

| **Total Reported Cases** | **New Reported Cases** | **Total Reported Deaths** | **New Reported Deaths** | 
| ---- | ---- | ---- | ---- |
| `r format(sum(cont_sums$cum_cases), big.mark=",")` | `r format(sum(cont_sums$cases), big.mark=",")` | `r format(sum(cont_sums$cum_deaths), big.mark=",")`|`r format(sum(cont_sums$deaths), big.mark=",")`|

<br>

```{r starter, echo = FALSE, collapse=TRUE, message = FALSE, warning = FALSE, fig.width=fig_height(7,9), fig.height=fig_height(4,5),fig.align=fig_center()}

suppressMessages(suppressWarnings(d %>% filter(date < date_0) %>% 
  group_by(date, Continent) %>% 
  summarise(deaths = sum(cum_deaths)) %>% 
  ggplot(aes(x=date,y=deaths,color=Continent)) + 
  geom_line(lwd=1) + 
  theme_bw() + xlab("") + ylab("Cumulative Deaths") + 
  scale_color_brewer(palette = "Set2") + 
  scale_y_log10(limits=c(1,NA), labels = scales::comma) + 
  theme(panel.border = element_blank(), axis.line = element_line(size=0.5))))

```

`r pagebreak()`

### Projected deaths within the next 28 days

<br>

```{r deaths plot, echo = FALSE, collapse=TRUE, message = FALSE, warning = FALSE,  fig.cap=figRef("**Figure 1: Cumulative Deaths since 10 deaths**. The projected deaths (assuming population mobility is maintained at today's level) are shown with dashed lines."), fig.width=fig_height(7,9), fig.height=fig_height(5,6),fig.align=fig_center()}

# [FIGURE SHOWING TOTAL DEATHS AND TOTAL CASES [OBSERVED AND INFERRED FOR SUB-SAHARAN AFRICA]
suppressMessages(one_firework_plot(cont = continent))

```

<br>
`r pagebreak()`

### Projected healthcare demand within the next 28 days

<br>

```{r healthcare plots, echo = FALSE, collapse=TRUE, message = FALSE, warning = FALSE, fig.cap=figRef("**Figure 2: Estimated Healthcare demands in 4 weeks time.** Individuals needing an ICU bed are predicted to need mechanical ventilation. Individuals requiring oxygen are assumed to require high-flow oxygen."), fig.width=fig_height(9,9), fig.height=fig_height(9,9),fig.align=fig_center()}

# [FIGURE SHOWING PROJECTIONS OF HOSP/ICU/DEATHS IN NEXT TWO WEEKS IN SSA]
suppressMessages(summaries_forecasts_plot(sums = summaries, cont = continent))

```


<br>
`r pagebreak()`

### Effective Reproductive Number

<br>

```{r rt plot, echo = FALSE, collapse=TRUE, message = FALSE, warning = FALSE,  fig.cap=figRef("**Figure 3: Time-varying reproduction number, $R_t$.** Dark green shows the 50% CI and light green shows the 95% CI. $R_t$ < 1 indicates a slowing epidemic, whereas an $R_t$ greater than 1 indicates a growing epidemic."), fig.width=fig_height(10,10), fig.height=fig_height(10,10),fig.align=fig_center()}

# [FIGURE SHOWING TOTAL DEATHS AND TOTAL CASES [OBSERVED AND INFERRED FOR SUB-SAHARAN AFRICA]
suppressMessages(rt_plot(cont = continent))

```

<br>
