---
title: ""
header-includes: \usepackage{caption}
date: ""
urlcolor: blue
output:
  html_document:
    keep_md: yes
    self_contained: yes
    css: styles.css
    theme: cosmo
    fig_caption: TRUE
    includes:
      in_header: ganalytics.txt
  pdf_document: 
    fig_caption: true
    toc: false
    includes: 
      in_header: header.tex
      after_body: footer.tex
  word_document:
    df_print: kable
    fig_caption: yes
    fig_height: 4
    fig_width: 6
    highlight: tango
    reference_docx: template.docx
params:
   o_list: NA
   replicates: 10
   data: NA
   date_0: NA
   country: NA
   surging: FALSE
   rt: NA
   rt_futures: NA
---

```{r rm_title_page, echo=FALSE}
head <- cat('
\\AtBeginDocument{\\let\\maketitle\\relax}
', file = "header.tex")

```

```{r knitr_options, echo=FALSE} 
knitr::opts_chunk$set(message = FALSE,warning = FALSE)
options("dplyr.summarise.inform"=FALSE)
```


```{js, echo = FALSE}

if ($(window).width() < 768) {
$('.dropdown-menu a.dropdown-toggle').on('click', function(e) {
if (!$(this).next().hasClass('show')) {
$(this).parents('.dropdown-menu').first().find('.show').removeClass("show");
}



var $subMenu = $(this).next(".dropdown-menu");
if (!$subMenu.hasClass('show')) {
$subMenu.addClass('show');
$subMenu.show();
} else {
$subMenu.removeClass('show');
$subMenu.hide();
}



$(this).parents('li.nav-item.dropdown.show').on('hidden.bs.dropdown', function(e) {
$('.dropdown-submenu .show').removeClass("show");
});



return false;
});
}

```


```{r, echo = FALSE, message=FALSE, warning=FALSE}
date <- as.Date(date)

# grab inputs
start <- 10
replicates <- params$replicates
data <- params$data 
country <- params$country
date_0 <- params$date_0
surging <- params$surging
rt <- params$rt
rt_futures <- params$rt_futures
url <- paste0("https://github.com/mrc-ide/global-lmic-reports/raw/master/",iso3c,"/index.pdf")

format_type <- knitr::opts_knit$get("rmarkdown.pandoc.to")
figRef <- function(caption) {
  if(format_type == "latex") {
    gsub("Figure \\d: ","",caption)
  } else {
    caption
  }
}

fig_height <- function(pdf = 4, html = 6) {
  if(format_type == "latex") {
    pdf
  } else {
    html
  }
}

fig_center <- function() {
  if(format_type == "latex") {
    "default"
  } else {
    "center"
  }
}

latex_omit <- function(text) {
  if(knitr::is_latex_output())
    return("")
  else
    return(text)
}

pagebreak <- function() {
  if(knitr::is_latex_output())
    return("\\newpage")
  else
    return('<div style="page-break-before: always;" />')
}

linebreak <- function() {
  if(knitr::is_latex_output())
    return("")
  else
    return('<br>')
}

pdf_line_break <- function() {
  if(knitr::is_latex_output())
    return("<br>")
  else
    return('')
}

if(data$deaths[1] >= 10) {
  show_sentence <- "" 
} else  {
  show_sentence <-  paste("**N.B. ", country, 
                          " is not shown in the following plot as only ", 
                          data$deaths[1],
                          " deaths have been reported to date**")
}

if(any(data$daily_cases < 0)) {
  show_cases_sentence <- paste("**N.B. ", country, 
                               " has revised their historic reported cases and", 
                               " thus have reported negative cases.**")
} else  {
  show_cases_sentence <- ""
}


```

```{r ecdc prep, echo = FALSE, collapse=TRUE, warning=FALSE, message=FALSE}

## Summaries
form <- function(x) {
  paste0(base::format(round(x[[1]]), big.mark=","),
         " (95% CI: ", base::format(round(x[[2]]), big.mark=","),
         "-", base::format(round(x[[3]]), big.mark=","),")")
}

form_round_dig <- function(x, dig = 2) {
  paste0(base::format(round(x[[1]], dig), big.mark=","),
         " (95% CI: ", base::format(round(x[[2]], dig), big.mark=","),
         "-", base::format(round(x[[3]], dig), big.mark=","),")")
}

t_test_safe <- function(x, ...) {
  out <- try(t.test(x, ...), silent = TRUE)
  if (inherits(out, "try-error"))
  {
    out <- list("conf.int"=c(mean(x),mean(x)))
  }
  return(out)
}


## -----------------------------------------------------------------------------
## totals unmitigated
## -----------------------------------------------------------------------------

infections <- params$o_list[[1]][params$o_list[[1]]$compartment == "infections",]
infections_28 <- infections[infections$t <= 0 & infections$t > -28, ]
infs <- group_by(infections_28, replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(inf_tot = mean(tot, na.rm = TRUE), 
            inf_min = t_test_safe(tot)$conf.int[1],
            inf_max = t_test_safe(tot)$conf.int[2])

infs_today <- group_by(infections[infections$t==0,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

infs_28 <- group_by(infections[infections$t==28,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

hospital <- params$o_list[[1]][params$o_list[[1]]$compartment == "hospital_demand",]
hosp_today <- group_by(hospital[hospital$t==0,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

hosp_28 <- group_by(hospital[hospital$t==28,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

icu <- params$o_list[[1]][params$o_list[[1]]$compartment == "ICU_demand",]
icu_today <- group_by(icu[icu$t==0,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

icu_28 <- group_by(icu[icu$t==28,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

deaths <- params$o_list[[1]][params$o_list[[1]]$compartment == "deaths",]
deaths_28 <- group_by(deaths[deaths$t==28,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

## -----------------------------------------------------------------------------
## totals mitigated
## -----------------------------------------------------------------------------

infections_mit <- params$o_list[[2]][params$o_list[[2]]$compartment == "infections",]
infections_mit_28 <- infections_mit[infections_mit$t <= 0 & infections_mit$t > -28, ]
infs_mit <- group_by(infections_mit_28, replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(inf_tot = mean(tot, na.rm = TRUE), 
            inf_min = t_test_safe(tot)$conf.int[1],
            inf_max = t_test_safe(tot)$conf.int[2])

infs_today_mit <- group_by(infections_mit[infections_mit$t==0,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

infs_28_mit <- group_by(infections_mit[infections_mit$t==28,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

hospital_mit <- params$o_list[[2]][params$o_list[[2]]$compartment == "hospital_demand",]
hosp_today_mit <- group_by(hospital_mit[hospital_mit$t==0,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

hosp_28_mit <- group_by(hospital_mit[hospital_mit$t==28,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

icu_mit <- params$o_list[[2]][params$o_list[[2]]$compartment == "ICU_demand",]
icu_today_mit <- group_by(icu_mit[icu_mit$t==0,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

icu_28_mit <- group_by(icu_mit[icu_mit$t==28,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

deaths_mit <- params$o_list[[2]][params$o_list[[2]]$compartment == "deaths",]
deaths_28_mit <- group_by(deaths_mit[deaths_mit$t==28,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

## -----------------------------------------------------------------------------
## totals reverse
## -----------------------------------------------------------------------------

infections_rev <- params$o_list[[3]][params$o_list[[3]]$compartment == "infections",]
infections_rev_28 <- infections_rev[infections_rev$t <= 0 & infections_rev$t > -28, ]
infs_rev_28 <- group_by(infections_rev, replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(inf_tot = mean(tot, na.rm = TRUE), 
            inf_min = t_test_safe(tot)$conf.int[1],
            inf_max = t_test_safe(tot)$conf.int[2])

infs_28_rev <- group_by(infections_rev[infections_rev$t==28,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

infs_today_rev <- group_by(infections_rev[infections_rev$t==0,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])


hospital_rev <- params$o_list[[3]][params$o_list[[3]]$compartment == "hospital_demand",]
hosp_today_rev <- group_by(hospital_rev[hospital_rev$t==0,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

hosp_28_rev <- group_by(hospital_rev[hospital_rev$t==28,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

icu_rev <- params$o_list[[3]][params$o_list[[3]]$compartment == "ICU_demand",]
icu_today_rev <- group_by(icu_rev[icu_rev$t==0,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

icu_28_rev <- group_by(icu_rev[icu_rev$t==28,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

deaths_rev <- params$o_list[[3]][params$o_list[[3]]$compartment == "deaths",]
deaths_28_rev <- group_by(deaths_rev[deaths_rev$t==28,], replicate) %>% 
  summarise(tot = sum(y, na.rm = TRUE)) %>% 
  summarise(i_tot = mean(tot, na.rm = TRUE), 
            i_min = t_test_safe(tot)$conf.int[1],
            i_max = t_test_safe(tot)$conf.int[2])

summary_number_df <- data.frame(variable = c("est_infections","report_infections",
                                             "report_deaths", "hospital_28", "icu_28",
                                             "hospital_28_mit", "icu_28_mit",
                                             "hospital_28_rev", "icu_28_rev"),
                                value = c(infs[[1]],  data$deaths[1], data$cases[1], 
                                          hosp_28[[1]], icu_28[[1]], 
                                          hosp_28_mit[[1]],icu_28_mit[[1]],
                                          hosp_28_rev[[1]],icu_28_rev[[1]]),
                                country = country,
                                continent = countrycode::countrycode(country, origin = 'country.name', destination = 'continent'))
saveRDS(summary_number_df, "summary_df.rds")

```

```{r capacity warn, echo = FALSE, collapse=TRUE, warning=FALSE}
if (surging) {
  show_capacity_sentence <- paste(
    "**N.B.", country,
    "is forecast to be close to or surpassing our best estimates for healthcare",
    "capacity in the next 28 days.** Estimates of deaths in the next 28 days may be", 
    "inaccurate due to our working assumptions for mortality in individuals who", 
    "do not receive appropriate treatment. [See our methods for more", 
    "information.](https://mrc-ide.github.io/global-lmic-reports/parameters.html)")
  fig_cap_capacity <- paste(
    "The forecasted deaths in blue assumes healthcare capacity has been surged",
    "to ensure sufficient supply of ICU and hospital beds. The red curve assumes",
    "no surging in healthcare capacity and subsequently projects increased deaths.")
} else  {
  show_capacity_sentence <-  paste("")
  fig_cap_capacity <- ""
}
```

```{r other sources, echo = FALSE, collapse = TRUE, warning = FALSE}

if (iso3c == "BRA") {
  other_sources <- paste("**For sub-national estimates of $R_t$, and further analysis of",
                         "Brazil, please see ",
                         "[Report 21](https://www.imperial.ac.uk/mrc-global-infectious-disease-analysis/covid-19/report-21-brazil/)**")
} else if (iso3c == "SYR") {
  other_sources <- paste("**We are aware of under-reporting of deaths in Damascus, Syria. This is not represented in this",
                         "report, but please see ",
                         "[Report 31](https://www.imperial.ac.uk/mrc-global-infectious-disease-analysis/covid-19/report-31-syria/)**")
} else if (iso3c == "SDN") {
  other_sources <- paste("**We are aware of under-reporting of deaths in Khartoum, Sudan. This is not represented in this",
                         "report, but please see ",
                         "[Report 39](https://www.imperial.ac.uk/mrc-global-infectious-disease-analysis/covid-19/report-39-sudan/)**")
} else {
  other_sources <- ""
}

```

## Situation Report for COVID-19: `r country`, `r base::format(date, "%Y-%m-%d")`

`r latex_omit("---")`

#### **[Download the report for `r country`, `r base::format(date, "%Y-%m-%d")` here](`r url`).**

This report uses data from COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University. These data are updated daily and whilst there may be a short delay, they are generally consistent with Ministry reports. These data are then used to back-calculate an ‘inferred number of COVID-19 infections’ using mathematical modelling techniques (see [Report 12](https://www.imperial.ac.uk/mrc-global-infectious-disease-analysis/covid-19/report-12-global-impact-covid-19/) for further details) to estimate the number of people that have been infected and to make short-term projections for future healthcare needs.

<br> 

### Epidemiological Situation

| **Total Reported Cases** | **New Reported Cases** | **Total Reported Deaths** | **New Reported Deaths** | **Estimated $R_{eff}$** |
| ---- | ---- | ---- | ---- | ------ |
| `r base::format(data$cases[1], big.mark=",")` | `r base::format(data$daily_cases[1], big.mark=",")` | `r base::format(data$deaths[1], big.mark=",")`|`r base::format(data$daily_deaths[1], big.mark=",")`| `r form_round_dig(rt$rts[nrow(rt$rts),c("Reff", "Reff_min", "Reff_max")])` |

`r pdf_line_break()`

The figure below shows the cumulative reported deaths as a function of the time since the 10th death was reported. Dashed lines show the expected trajectory for different doubling times of the epidemic. For example, with a doubling time of 3 days, if there are currently a total of 20 deaths reported, we would expect there to be 40 deaths in total reported in 3 days-time, 80 deaths in 6 days-time, 160 deaths in 9 days-time etc. For most epidemics, in the absence of interventions, we expect a doubling time of 3-4 days for this disease. `r show_sentence`

<br>

```{r fig1, echo = FALSE, collapse=TRUE, message = FALSE, warning = FALSE, fig.cap=figRef("**Figure 1: Cumulative Deaths since 10 deaths.** Country not shown if fewer than 10 deaths."), fig.height=fig_height(4.25,6), fig.width=fig_height(8,8), fig.align=fig_center()}

suppressWarnings(suppressMessages(print(cumulative_deaths_plot(country = country))))

```

\newpage

<br>

### COVID-19 Transmission Modelling

The figure below shows the estimated number of people infected over the past 4 weeks. The bar charts show, for comparison, the number of reported cases. We estimate that there has been a total of `r form(infs)` infections over the past 4 weeks. The right-hand plot shows these data on a different scale as the estimated infections are likely to be much larger than the reported cases.  **Importantly**, the estimated infections includes both asymptomatic and mild cases that would not necessarily be identified through surveillance. Consequently, the estimated infections are likely to be significantly higher than the reported cases in all countries (see our [FAQ](https://mrc-ide.github.io/global-lmic-reports/FAQ.html#why-are-the-estimated-infections-so-much-higher-than-the-reported-cases) for further explanation of these differences and why the reported cases and estimated infections are unlikely to match). `r show_cases_sentence`

<br>

```{r case plot, echo = FALSE, collapse=TRUE, message = FALSE, warning = FALSE,  fig.cap=figRef("**Figure 2: Daily number of infections estimated by fitting to the current total of deaths.** Reported cases are shown in red. Model estimated infections are shown in blue (dark blue 50% interquartile range, light blue 95% quantile). The dashed line shows the current day."), fig.width=10, fig.height=fig_height(5.5,5.5), fig.align=fig_center()}

suppressWarnings(suppressMessages(print(cases_plot(params$o_list[[1]], data = data, date = date, date_0 = date_0))))

```

`r pagebreak()`

By fitting to the time series of deaths, we are able to estimate a time-varying reproduction number, $R_{eff}$. $R_{eff}$ is the the average number of secondary infections caused by a single infected person at a given time. If $R_{eff}$ is above 1, the rate of transmission is increasing and the number of new infections is increasing. $R_{eff}$ is assumed to change in relation to mobility  fall in proportion. When fitting our model we assume that 100% of COVID-19 related deaths have been reported (please see our [FAQ](https://mrc-ide.github.io/global-lmic-reports/FAQ.html#how-have-you-accounted-for-unreported-deaths) section for more information about this assumption). Additionally, we assume that infection with COVID-19 leads to protective immunity that does not wane within the time scales considered in these analyses.

<br>

`r other_sources`

<br>

```{r fig0, echo = FALSE, collapse=TRUE, message = FALSE, warning = FALSE, fig.cap=figRef("**Figure 3: Time-varying effective reproduction number, $R_{eff}$.** $R_{eff}$ (**green**) is the average number of secondary infections caused by a single infected person at time equal to $t$. A horizonatal dashed line is shown at $R_{eff}$ = 1. $R_{eff}$ < 1 indicates a slowing epidemic in which new infections are not increasing. $R_{eff}$ > 1 indicates a growing epidemic in which new infections are increasing over time. Dark green shows the 50% CI and light green shows the 95% CI"), fig.height=fig_height(3.5,5), fig.width=fig_height(6,10), fig.align=fig_center()}

suppressWarnings(suppressMessages(print(rt$plot)))

```

<br>

`r pagebreak()`

Using the model fit, we can forecast the expected trajectory for cumulative deaths assuming the transmission level, represented by the final $R_{t}$ value stays the same over the next 28 days. `r show_capacity_sentence`

<br>

```{r death forecast plots, echo = FALSE, collapse=TRUE, warning=FALSE, message=FALSE, fig.cap=figRef(paste0("**Figure 4: Estimated daily deaths.** Projected deaths assuming the current level of interventions are maintained are shown in red (median and 95% quantile). Reported deaths are plotted in black. The plot on the left is focussed on the model fit prior to today, while the plot on the right forecasts the next 28 days. ", fig_cap_capacity)), fig.width=if (surging) 10 else 8, fig.height=if (surging) 8 else 6, fig.align=fig_center()}

if (surging) {
  suppressWarnings(suppressMessages(
    deaths_plot_single_surge_vaccine(params$o_list[[1]], params$o_list[[4]],
                             forecast = 28, 
                       data = data, date_0 = date_0, date = date)))  
} else {
  suppressWarnings(suppressMessages(print(
    deaths_plot_single_vaccine(params$o_list[[1]], forecast = 28, 
                       data = data, date_0 = date_0, date = date))))
  
}


```

<br>

`r pagebreak()`

### Short-term Epidemic Scenarios

```{r, include=FALSE, echo=FALSE}
rt_futures <- rt_futures %>% 
  filter(!surged)
optimisitic <- paste0(signif(rt_futures$optimistic_med, 3), " (", 
                      signif(rt_futures$optimistic_025, 3), ", ",
                      signif(rt_futures$optimistic_975, 3), ")")

pessimistic <- paste0(signif(rt_futures$pessimistic_med, 3), " (", 
                      signif(rt_futures$pessimistic_025, 3), ", ",
                      signif(rt_futures$pessimistic_975, 3), ")")
```


We make the following short-term projections of healthcare demand and new infections under the following three scenarios:

* **Scenario 1.** The epidemic continues to grow at the current rate. 
* **Scenario 2.** The $R_{t}$ value will decrease to `r optimisitic`, representing
either the implementation of new interventions, scaling up of interventions, or
a reduction in mobility.
* **Scenario 3.** The $R_{t}$ value will increase to `r pessimistic`, representing
either the relaxation of interventions, or an increase in mobility.

The $R_{t}$ values for scenarios 2 and 3 are calculated by estimating the linear 
trends in $R_{t}$ post-lockdown across 14-day periods, for each replicate from the MCMC, 
and then finding the 90% percentile of the positive and negative gradients.
These are then applied to the final fitted $R_{t}$ value, for each replicate.
Hence the optimistic and and pessimistic scenarios represent if the upper limit of historic 
negative or positive trends in $R_{t}$ are applied from this time on wards, including both
changes in $R_{t}$ due to mobility and non-mobility based interventions.

We estimate that over the next 4 weeks demand for hospital beds will change from `r form(hosp_today)` patients requiring treatment with high-pressure oxygen at the current date to `r form(hosp_28)` hospital beds being required on `r base::format(date + 28, "%Y-%m-%d")` if no further interventions are introduced (Scenario 1). Similarly, we estimate that over the next 4 weeks demand for critical care (ICU) beds will change from `r form(icu_today)` patients requiring treatment with mechanical ventilation at the current date to `r form(icu_28)` by `r base::format(date + 28, "%Y-%m-%d")`. These projections assume that approximately 5% of all infections will require treatment with high-pressure oxygen and that approximately 30% of hospitalised cases will require treatment with mechanical ventilation (based on analysis of ongoing epidemics in Europe). **N.B. These scenarios are unlikely to show significant differences for the first week since there is a delay of approximately 10 days between infection and hospital admission. Consequently, the effectiveness of a change in policy is likely to be better captured by hospital admission data approximately 2 weeks after the policy change is implemented.**

<br>

```{r healthcare plots, echo = FALSE, collapse=TRUE, message = FALSE, warning = FALSE, fig.cap=figRef("**Figure 5: Healthcare demands in the next 28 days.** Individuals needing an ICU bed are assumed to need mechanical ventilation. Projected demand for Scenario 1 (the epidemic continues to grow at the current rate) are shown in green (Maintain status quo). Projections for Scenario 2 (optimistic) are shown in blue. Projections for Scenario 3 (pessimistic) are shown in red. Current date shown with dashed line."), fig.width=fig_height(8,10), fig.height=fig_height(3.5,4.5), fig.align=fig_center()}

# suppressWarnings(suppressMessages(cowplot::plot_grid(
#   healthcare_plot_contrast(o1 = params$o_list[[1]], o2 = params$o_list[[2]], 
#                            data = data, date_0 = date_0, date = date, 
#                            forecast = 28, what = "hospital_demand"),
#   healthcare_plot_contrast(o1 = params$o_list[[1]], o2 = params$o_list[[2]], 
#                            data = data, date_0 = date_0, date = date, 
#                            forecast = 28, what = "ICU_demand"),
#                    ncol=2)))

suppressWarnings(suppressMessages(gg_hosp <- healthcare_plot_contrast_triple(o1 = params$o_list[[1]], 
                                           o2 = params$o_list[[2]], 
                                           o3 = params$o_list[[3]], 
                                           data = data, date_0 = date_0, date = date, 
                                           forecast = 28, what = "hospital_demand")))
suppressWarnings(suppressMessages(gg_icu <- healthcare_plot_contrast_triple(o1 = params$o_list[[1]], 
                                          o2 = params$o_list[[2]], 
                                          o3 = params$o_list[[3]], 
                                          data = data, date_0 = date_0, date = date, 
                                          forecast = 28, what = "ICU_demand")))
leg <- suppressWarnings(suppressMessages(
  cowplot::get_legend(gg_hosp + theme(legend.justification = c(0.5,1)))
))
bottom <- suppressWarnings(suppressMessages(
  cowplot::plot_grid(gg_hosp + theme(legend.position = "none"), 
                     gg_icu + theme(legend.position = "none"), ncol = 2)
))
suppressWarnings(suppressMessages(grid <- cowplot::plot_grid(leg,bottom,nrow=2,rel_heights = c(0.1,1))))
suppressWarnings(suppressMessages(print(grid)))

```

<br>
`r pagebreak()`

The impact of each scenario has a more immediate effect on the daily number of infections. The figure below shows the impact of each scenario on the estimated daily incidence of new infections. If interventions are scaled up (Scenario 2), the daily number of infections will change from `r form(infs_today)` at the current date to `r form(infs_28_mit)` by `r base::format(date + 28, "%Y-%m-%d")`. Under the pessimistic scenario, we estimate the daily number of infections will change from `r form(infs_today)` at the current date to `r form(infs_28_rev)` by `r base::format(date + 28, "%Y-%m-%d")`.

<br>

```{r infections forecast plots, echo = FALSE, collapse=TRUE, warning=FALSE, message=FALSE, fig.cap=figRef("**Figure 6: Daily number of infections estimated by fitting to deaths.** Projected infections for Scenario 1 (the epidemic continues to grow at the current rate) are shown in green (Maintain status quo). Projections for Scenario 2 (optimistic) are shown in blue. Projections for Scenario 3 (pessimistic) are shown in red. Current date shown with dashed line."), fig.width=7, fig.height=fig_height(4,4.5), fig.align=fig_center()}


suppressWarnings(suppressMessages(print(cases_contrast_triple_bars(
  o1 = params$o_list[[1]], 
  o2 = params$o_list[[2]], 
  o3 = params$o_list[[3]], 
  forecast = 28,
  data = data, date_0 = date_0, date = date))))

```

---

To explore different scenarios, we recommend using our [COVID-19 Scenario Analysis Tool - https://covidsim.org/](https://covidsim.org/), which can be used to simulate different intervention scenarios and explore the long term impact on healthcare demand. 

---
