---
title: News
header-includes: \usepackage{caption}
date: ""
output:
  html_document:
    keep_md: yes
    self_contained: yes
    css: styles.css
    theme: cosmo
    fig_caption: TRUE
    includes:
      in_header: ganalytics.html
  pdf_document: 
    fig_caption: true
    includes: 
      before_body: footer.tex
  word_document:
    df_print: kable
    fig_caption: yes
    fig_height: 4
    fig_width: 6
    highlight: tango
    reference_docx: template.docx
---

```{js, echo = FALSE}

if ($(window).width() < 768) {
  $('.dropdown-menu a.dropdown-toggle').on('click', function(e) {
    if (!$(this).next().hasClass('show')) {
      $(this).parents('.dropdown-menu').first().find('.show').removeClass("show");
    }

 

    var $subMenu = $(this).next(".dropdown-menu");
    if (!$subMenu.hasClass('show')) {
      $subMenu.addClass('show');
      $subMenu.show();
    } else {
      $subMenu.removeClass('show');
      $subMenu.hide();
    }

 

    $(this).parents('li.nav-item.dropdown.show').on('hidden.bs.dropdown', function(e) {
      $('.dropdown-submenu .show').removeClass("show");
    });

 

    return false;
  });
}

```

```{r, echo = FALSE, message=FALSE, warning=FALSE}
start <- 10
replicates <- 10
library(dplyr)
library(ggplot2)
library(squire)

format <- knitr::opts_knit$get("rmarkdown.pandoc.to")
figRef <- function(caption) {
  if(format == "latex") {
    gsub("Figure \\d: ","",caption)
  } else {
    caption
  }
}


fig_height <- function(pdf = 4, html = 6) {
    if(format == "latex") {
    pdf
  } else {
    html
  }
}

fig_center <- function() {
    if(format == "latex") {
    "default"
  } else {
    "center"
  }
}


```

```{r, echo = FALSE, message=FALSE, warning=FALSE}
render_toc <- function(
  filename, 
  toc_header_name = "Table of Contents",
  base_level = NULL,
  toc_depth = 3,
  section = "Questions"
) {
  x <- readLines(filename, warn = FALSE)
  x <- paste(x, collapse = "\n")
  x <- paste0("\n", x, "\n")
  for (i in 5:3) {
    regex_code_fence <- paste0("\n[`]{", i, "}.+?[`]{", i, "}\n")
    x <- gsub(regex_code_fence, "", x)
  }
  x <- strsplit(x, "\n")[[1]]
  x <- x[grepl("^#+", x)]
  if (!is.null(toc_header_name)) 
    x <- x[!grepl(paste0("^#+ ", toc_header_name), x)]
  if (is.null(base_level))
    base_level <- min(sapply(gsub("(#+).+", "\\1", x), nchar))
  start_at_base_level <- FALSE
  x <- sapply(x, function(h) {
    level <- nchar(gsub("(#+).+", "\\1", h)) - base_level
    if (level < 0) {
      stop("Cannot have negative header levels. Problematic header \"", h, '" ',
           "was considered level ", level, ". Please adjust `base_level`.")
    }
    if (level > toc_depth - 1) return("")
    if (!start_at_base_level && level == 0) start_at_base_level <<- TRUE
    if (!start_at_base_level) return("")
    if (grepl("\\{#.+\\}(\\s+)?$", h)) {
      # has special header slug
      header_text <- gsub("#+ (.+)\\s+?\\{.+$", "\\1", h)
      header_slug <- gsub(".+\\{\\s?#([-_.a-zA-Z]+).+", "\\1", h)
    } else {
      header_text <- gsub("#+\\s+?", "", h)
      header_text <- gsub("\\s+?\\{.+\\}\\s*$", "", header_text) # strip { .tabset ... }
      header_text <- gsub("^[^[:alpha:]]*\\s*", "", header_text) # remove up to first alpha char
      header_slug <- paste(strsplit(header_text, " ")[[1]], collapse="-")
      header_slug <- tolower(header_slug)
    }
    paste0(strrep(" ", level * 4), "- [", header_text, "](#", header_slug, ")")
  })
  x <- x[x != ""]
  mains <- grep("^-",x)
  for(i in seq_len(length(x))) {
    x[i] <- gsub("- ", paste0(i-1,". "), x[i])
    x[i] <- gsub("\\?)$",")", x[i])
    x[i] <- gsub("?-/-","-",x[i],fixed=TRUE)
  }
  x <- stringr::str_trim(x[-mains])
  #x <- gsub("^\\d\\. ","", x)
  
  knitr::asis_output(paste(rev(x), collapse = " \n"))
}
```


## News

SARS-CoV-2 is a novel virus and we gain new information on its transmission almost daily, with an incredible amount of new research on COVID-19 being published each week. Consequently, as our understanding of COVID-19 changes over time, we aim to respond accordingly and update our predictions. 

Here, we detail the major version changes to the reports produced in this website (minor changes, such as typo fixes and small edits to plots and text are not documented here). These major changes represent either a new methodology or new information being presented. We hope the following overview will help users understand the changes implemented throughout the pandemic. 

Lastly, as we switch from one version to the next, there is likely to be a small delay in keeping the website up to date. We apologise for the delays.

---

#### Version 4: 2020-06-28

* Mobility independent change in transmission is now also modelled (alongside mobility dependent transmission) post lockdown to reflect change in people's behaviour that is not captured by mobility. 

#### Version 3: 2020-06-14

* Impact of mobility on transmision estimated both pre and post lockdown for each country, allowing mobility and transmission to be able to decouple after lockdowns are relaxed if the data supports this. 

#### Version 2: 2020-05-31
  
* Data on mobility from the [Google Community Mobility Reports](https://www.google.com/covid19/mobility/) to inform changes in transmission.
* Estimates of $R_t$ included in country reports
* Projections extended to 28 days
* Impact of countries nearing health system capacity explored
* Regional overviews provided
* Surged healthcare capacity explored (2020-06-02)
* Fix for looking at healthcare capacity being surged (2020-06-09)
* Scenario projections provided in zipped files and update historically for each of pipelines (2020-06-09)

<br>

#### Version 1: 2020-04-23

* Initial version released, with reports presenting information on the following:
  - The total number of COVID-19 infections
  - The expected number of deaths within the next 14 days
  - The number of individuals requiring oxygen or mechanical ventilation in the next 14 days
  - The impact of changing their current intervention policy
* Model outputs available to download from home page


<br>